### 2008-08-10 papers/OTSPDE-Systems/OTSPDE-Systems.tex
OTSPDE: * Checked in place holders for upcoming papers. * Updated information on potential journals to submit to.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Manuscript started: 2008/09/0?
% First completed draft: 2008/??/??, Vitamin D, Inc.
% Submission draft: 2008/??/??, Vitamin D, Inc.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[oneeqnum,onefignum,onetabnum,onethmnum]{siamltex}
\usepackage{graphics}
\usepackage{multirow}

\include{macros}

\title{High-Order Accurate Finite Difference Schemes \\
       for Systems of PDEs via Optimal Time Step Selection}

\author{
Kevin T. Chu\footnotemark[2] \footnotemark[3]
}

\begin{document}
\bibliographystyle{unsrt}
\maketitle

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[2]{Vitamin D, Inc., Menlo Park, CA 94025}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[3]{Institute of High Performance Computing, A*STAR, Singapore, Singapore}

\renewcommand{\thefootnote}{\arabic{footnote}}

\begin{abstract}
place holder...
\end{abstract}

\begin{keywords}
optimal time step, finite difference schemes, high-order accurate numerical 
methods, time dependent PDEs, systems of PDEs
\end{keywords}

\begin{AMS}
65-02, 65M06, 65M12, 65M20, ??
\end{AMS}

\pagestyle{myheadings}
\thispagestyle{plain}
\markboth{KEVIN T. CHU}
         {HIGH-ORDER FD SCHEMES VIA OPTIMAL TIME STEPS}

\section*{Introduction}
High-order numerical methods for partial differential equations (PDEs) will 
always be valuable for increasing the computational efficiency of numerical 
simulations.  Thus, it is not at all surprising that a great deal of effort in 
numerical PDEs continues to be focused on the development of high-order 
numerical schemes~\cite{bruger_2005, gibou_2005, ito_2005, shukla_2005, 
shukla_2007}.  
Typically, high-order accuracy is achieved by constructing
schemes that have high formal orders of accuracy.  However, high-order 
accurate numerical solutions can also be obtained by using formally low-order 
schemes in clever ways.  When possible, the latter approach can be a powerful 
way to boost the accuracy of a numerical method \emph{without} introducing too 
much additional algorithmic (and programming) complexity.

\section{\label{sec:summary} Summary}

\section*{Acknowledgments}
The author gratefully acknowledges the support of Vitamin D, Inc.
and the Institute for High-Performance Computing (IHPC) in Singapore. 
The author would like to thank K.-H. Chiam, ?? 
for helpful suggestions on the manuscript.  
The author particularly thanks K.-H. Chiam for pointing him to the 
mathematical biology literature for interesting examples of reaction-diffusion 
systems.

\bibliography{OTSPDE-Systems}

\end{document}