### 2008-08-10 papers/OTSPDE-VarCoefs/OTSPDE-VarCoefs.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Manuscript started: 2008/09/0?
% First completed draft: 2008/??/??, Vitamin D, Inc.
% Submission draft: 2008/??/??, Vitamin D, Inc.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[oneeqnum,onefignum,onetabnum,onethmnum]{siamltex}
\usepackage{graphics}
\usepackage{multirow}

\include{macros}

\title{High-Order Accurate Finite Difference Schemes \\
       for Variable-Coefficient PDEs via Optimal Time Step Selection}

\author{
Kevin T. Chu\footnotemark[2] \footnotemark[3]
and 
James V. Lambers\footnotemark[4]
}

\begin{document}
\bibliographystyle{unsrt}
\maketitle

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[2]{Vitamin D, Inc., Menlo Park, CA 94025}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[3]{Institute of High Performance Computing, A*STAR, Singapore, Singapore}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[4]{Department of Energy Resources Engineering, Stanford University, Stanford, CA 94305}

\renewcommand{\thefootnote}{\arabic{footnote}}

\begin{abstract}
place holder
\end{abstract}

\begin{keywords}
optimal time step, finite difference schemes, high-order accurate numerical 
methods, time dependent PDEs, variable-coefficient PDEs
\end{keywords}

\begin{AMS}
65-02, 65M06, 65M12, 65M20 ??
\end{AMS}

\pagestyle{myheadings}
\thispagestyle{plain}
\markboth{KEVIN T. CHU AND JAMES V. LAMBERS}
         {??}

\section*{Introduction}
\cite{zoppou1999, kreiss2002, guidotti2006}

% ACKNOWLEDGMENTS
\section*{Acknowledgments}
The author gratefully acknowledges the support of Vitamin D, Inc.
and the Institute for High-Performance Computing (IHPC) in Singapore. 
The author would like to thank ??
for helpful suggestions on the manuscript.

\bibliography{OTSPDE-VarCoefs}

\end{document}