### 2009-05-25 papers/OTSPDE/submissions/200905-Applied_Numerical_Math/cover_letter.tex

\documentclass[12pt]{letter}

% page settings
\textwidth 6.5in
\textheight 9in
\oddsidemargin -0.05in
\evensidemargin -0.05in
\topmargin -0.85in

\begin{document}

\address{}
\signature{Kevin T. Chu}

\begin{letter}
{
% NORMALLY ADDRESS GOES HERE
}

\opening{Dear Applied Numerical Mathematics Editor/Referee,}

My article titled \emph{Boosting the Accuracy of Finite Difference Schemes via
Optimal Time Step Selection and Non-Iterative Defect Correction} discusses a 
simple technique for boosting the order of accuracy of finite difference 
schemes for time dependent PDEs by choosing an optimal, \emph{fixed} time step
and adding defect correction terms in a \emph{non-iterative} manner.  As 
discussed in the paper, a careful choice of time step can dramatically 
reduce the numerical error of simple finite difference schemes and make 
it possible to achieve high-order accuracy using only formally low-order 
finite difference stencils and time integration schemes.

This paper describes the method for calculating the optimal time step and
defect correcction terms and explains the observed boost in accuracy using 
straightforward numerical analysis arguments.  The technique of optimal time 
step selection is applicable to several types of finite difference schemes
for linear and semilinear PDEs in any number of space dimensions and both 
regular and irregular domains.  Examples of each of these types of problems 
are discussed and analyzed in the paper.

Thank you for your review of the paper.  I look forward to hearing your 
comments and suggestions for improvement of the manuscript.

\closing{Sincerely,}

\end{letter}

\end{document}